The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2016-08-23 script/en/chapters/homomorphicencryption.tex
Corrected typos and overfull hbox; congruent index (no redundancy in wording; adjectives in lowercase in German; all normal words in lowercase in English)
CT Book ed 12: Major update -- many changes, corrections, and additions during last 4 months

% ..............................................................................
% Homomorphe Chiffren
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

\newpage
\hypertarget{homciph}{}
\chapter{Homomorphic Ciphers}
\label{Chapter_HomomorphicCiphers}
\index{Homomorphic ciphers}\index{Encryption!homomorphic}
(Martin Franz, January 2013)

@@

The most famous cryptosystem with homomorphic properties is the one by Paillier \cite{hc:Paillier}. First we will see how the Paillier key generation process works. After that, we will show that the Paillier cryptosystem indeed has homomorphic properties.

@@

\item A second application of homomorphic ciphers is ``Secure Multiparty Computation''. Here, two or more parties can compute any commonly known function. Each of the parties provides one or more of the inputs for the function to be computed. The goal of the secure computations is to keep all private inputs secret, while only the result of the function is revealed. The use of homomorphic encryption helps to perform these computations on encrypted data. However, since the Paillier encryption only allows to compute additions of encrypted values (and, e.g. no multiplications can be performed), a number of additional methods and techniques have to be applied. The Wikipedia page \cite{hc:SMC} offers a great start for reading more about this topic and more advanced techniques for Secure Multiparty Computation.

\item Furthermore it is expected, that homomorphic encryption will provide great advantages in the areas of ``Cloud Computing''. Using so called ``fully-homomorphic encryption'' \cite{hc:HomEnc} it will be possible to run large applications on external servers only on encrypted data. For this, necessarily one needs to be able to perform both arithmetic operations, the addition and the multiplication, on encrypted data (in contrast to Paillier encryption, which only allows performing additions). Such a crypto system was first presented in 2009 \cite{hc:Gentry2009}.
\end{enumerate}

@@

\subsection{CrypTool 2}

In CrypTool 2 you will find an implementation of the Paillier cryptosystem. Among the available components, there are components for key generation (Paillier Key Generator), an example for encryption and decryption with Paillier (called Paillier Text), as well as examples which illustrate the homomorphic properties of the cryptosystem (Paillier Addition, Paillier Blinding and Paillier Voting).

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.8]{figures/CT2-Paillier.png}
\caption{Paillier cryptosystem in CrypTool 2} 
\label{CT2-Paillier}
\end{center}
\end{figure}

@@

In JCrypTool there is an implementation (see Figure \ref{JCT-HomEnc}), which visualizes the homomorphic properties of various cryptosystem. For RSA and Paillier it shows, that multiplications, for RSA, and additions for Paillier, respectively, can be performed on encrypted values. For the fully-homomorphic cryptosystem by Gentry it is possible to perform both multiplications, as well as additions on encrypted values.

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.4]{figures/JCT-HomEnc.PNG}
\caption{Visualization of homomorphic properties in JCrypTool} 
\label{JCT-HomEnc}
\end{center}
\end{figure}

%------------------------------------------------------------------------------
\newpage
\begin{thebibliography}{99999}
\addcontentsline{toc}{section}{Bibliography}

\bibitem[Paillier1999]{hc:Paillier} \index{Paillier 1999}
   Pascal Paillier, \\
   {\em Public-Key Cryptosystems Based on Composite Degree Residuosity Classes},\\
	 Advances in Cryptology -- EUROCRYPT'99, 1999.

\bibitem[SMC]{hc:SMC} \index{SMC}
   Wikipedia, \\
   {\em Secure Multiparty Computation}.\\
   \url{http://en.wikipedia.org/wiki/Secure_multi-party_computation}

\bibitem[HomEnc]{hc:HomEnc}
   Wikipedia, \\
   {\em Homomorphic Encryption}\\
   \url{https://en.wikipedia.org/wiki/Homomorphic_encryption}\\
   {\em Homomorphism}\\
   \url{https://en.wikipedia.org/wiki/Homomorphism}

\bibitem[Gentry2009]{hc:Gentry2009} \index{Gentry 2009}
   Craig Gentry, \\
   {\em Fully Homomorphic Encryption Using Ideal Lattices},\\
	In the 41st ACM Symposium on Theory of Computing (STOC), 2009.

\end{thebibliography}  % German

>>>>

% $Id:
% ..............................................................................
%          H o m o m o r p h i c   C i p h e r s
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

\begin{bibunit}[babalpha] %% alpha: Chapter bibliography shows authors abbreviation

\newpage
% #### old naming: \hypertarget{Homomorphic Ciphers}{}
\hypertarget{Chapter_HomomorphicCiphers}{}
\chapter{Homomorphic Ciphers}
\label{Chapter_HomomorphicCiphers}
\index{homomorphic ciphers}\index{encryption!homomorphic}
(\hyperlink{author_Martin-Franz}{Martin Franz}, Jan 2013)

@@

The most famous cryptosystem with homomorphic properties is the one by Paillier \cite{Paillier1999}. First we will see how the Paillier key generation process works. After that, we will show that the Paillier cryptosystem indeed has homomorphic properties.

@@

\item A second application of homomorphic ciphers is ``Secure Multiparty Computation''. Here, two or more parties can compute any commonly known function. Each of the parties provides one or more of the inputs for the function to be computed. The goal of the secure computations is to keep all private inputs secret, while only the result of the function is revealed. The use of homomorphic encryption helps to perform these computations on encrypted data. However, since the Paillier encryption only allows to compute additions of encrypted values (and, e.g. no multiplications can be performed), a number of additional methods and techniques have to be applied. The Wikipedia page \cite{Wiki_SMC} offers a great start for reading more about this topic and more advanced techniques for Secure Multiparty Computation.

\item Furthermore it is expected, that homomorphic encryption will provide great advantages in the areas of ``Cloud Computing''. Using so called ``fully-homomorphic encryption'' \cite{Wiki_HomEnc} it will be possible to run large applications on external servers only on encrypted data. For this, necessarily one needs to be able to perform both arithmetic operations, the addition and the multiplication, on encrypted data (in contrast to Paillier encryption, which only allows performing additions). Such a crypto system was first presented in 2009 \cite{Gentry2009}.
\end{enumerate}

@@

\subsection{CrypTool~2}

In CrypTool~2\index{CT2} you will find an implementation of the Paillier cryptosystem. Among the available components, there are components for key generation (Paillier Key Generator), an example for encryption and decryption with Paillier (called Paillier Text), as well as examples which illustrate the homomorphic properties of the cryptosystem (Paillier Addition, Paillier Blinding and Paillier Voting).

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.8]{figures/CT2-Paillier.png}
	\caption{Paillier cryptosystem in CrypTool~2 (CT2)} 
\label{CT2-Paillier}
\end{center}
\end{figure}

@@

In JCrypTool\index{JCT} there is an implementation (see Figure \ref{JCT-HomEnc}), which visualizes the homomorphic properties of various cryptosystem. For RSA and Paillier it shows, that multiplications, for RSA, and additions for Paillier, respectively, can be performed on encrypted values. For the fully-homomorphic cryptosystem by Gentry it is possible to perform both multiplications, as well as additions on encrypted values.

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.4]{figures/JCT-HomEnc.PNG}
	\caption{Visualization of homomorphic properties in JCrypTool (JCT)} 
\label{JCT-HomEnc}
\end{center}
\end{figure}

%------------------------------------------------------------------------------
\putbib[../de/references]
\addcontentsline{toc}{section}{Bibliography}
\end{bibunit}

\noindent All links have been confirmed at July 15, 2016.

### 2016-08-23 script/en/chapters/introduction.tex
CT Book ed 12: Major update -- many changes, corrections, and additions during last 4 months

% $Id$
% ............................................................................
%      V O R W O R T  und  E I N F U E H R U N G (Zusammenspiel CT-Buch und CT-Programme) 
%      P R E F A C E  and  I N T R O D U C T I O N (Playing together of book and programs)
%
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

\clearpage\phantomsection
\addcontentsline{toc}{chapter}{Preface to the 12th Edition of the CrypTool Book xxxxxxxxxxxx}
\chapter*{Preface to the 12th Edition of the CrypTool Book xxxxxxxxxxxx}

@@

Starting in the year 2000 this book became part of the CrypTool 1
(CT1)\index{CrypTool 1} package in version 1.2.01. 
Since then it has been expanded and revised in almost every new version of CT1.

@@

The later editorial work in LaTeX added footnotes and cross linkages between
different sections, and some corrections.

%This edition of the CT book again updated some topics:
Compared to edition 11, in this edition
% completely updated the TeX sources of the document, and of course,
the content of the book was amended, corrected, and
updated with some topics, for instance:
\vspace{-5pt}
\begin{itemize}
%  \item the definitions of the strength of security functions
%        (chapter \ref{cm_Section_Security_Definitions}),
  \item the largest prime numbers (chap. \ref{search_for_very_big_primes}),
%        new factorization records (chap. \ref{RSA-768}),
%  \item progress in cryptanalysis of AES 
%        (chap. \ref{NeueAES-Analyse}) and
%  \item progress in cryptanalysis of hash algorithms 
%        (chap. \ref{collision-attacks-against-sha-1}) and
%  \item progress in ideas for new crypto methods (RSA successor) 
%        (chap. \ref{xxxxxxxxxBrute-force-gegen-Symmetr})\index{xxxxxxxxxx} and
  \item the list of movies or novels, in which cryptography or number theory 
        played a major role (see appendix \ref{s:appendix-movies}),
%       and where primes are used as hangers  
%        (see curiouses in \ref{HT-Quaint-curious-Primes-usage}),
  \item the overviews of all functions in
        \hyperlink{appendix-template-overview-CT2}{CrypTool 2 (CT2)},
        \hyperlink{appendix-function-overview-JCT}{JCrypTool (JCT)}, and
        \hyperlink{appendix-function-overview-CTO}{CrypTool-Online (CTO)}
        (see appendix),
  \item further SageMath scripts for cryptography, and the appendix
        \ref{s:appendix-using-sage} about using the computer algebra system
        SageMath,
        % Nguyen and Massierer (see also chap.~\ref{ec:Sage_Massierer}).
  \item the section about the Goldbach conjecture
        (see \ref{L-GoldbachConjecture}) and about twin primes
        (see \ref{L-TwinCousinPrimes}),
  \item the section about shared primes in RSA modules used in reality
        (see \ref{nt_Shared-Primes}),
%  \item the section about RSA fixed points
%        (see \ref{l:NumberTheory_Sage_Number-of-RSA-FixedPoints}), and
%  \item homomorphic encryption
%        (see chapter \ref{Chapter_HomomorphicCiphers}).
  \item the \nameref{Chapter_BitCiphers} is completely new
        (see chapter \ref{Chapter_BitCiphers}).
\end{itemize}

@@

At this point I'd like to thank explicitly the following people who
particularly contributed to the CrypTool projct\index{CrypTool}. They
applied their very special talents and showed really great engagement:
\vspace{-7pt}
%\begin{itemize}
\begin{list}{\textbullet}{\addtolength{\itemsep}{-0.5\baselineskip}}
   \item Mr.\ Henrik Koy
   \item Mr.\ J\"org-Cornelius Schneider
   \item Mr.\ Florian Marchal
   \item Dr.\ Peer Wichmann
   \item Hr.\ Dominik Schadow
   \item Staff of 
         Prof.\ Johannes Buchmann,
         Prof.\ Claudia Eckert,
         Prof.\ Alexander May,
         Prof.\ Torben Weis and especially
         Prof.\ Arno Wacker.
\end{list}

@@

Thanks also to the readers who sent us feedback.
% Especially I would like to acknowledge the English language proof-reading
% of this book version done by Richard Christensen and Lowell Montgomery.

@@

\par \vskip + 45pt
\noindent Bernhard Esslinger
\par \vskip + 10pt
\noindent Frankfurt/Siegen (Germany), April 2016

\par \vskip + 100pt
\noindent PS:\\
We'd be glad if new authors would show up to add further chapters,
e.g. about\\
- the Riemann Zeta function,\\
- hash functions,\\
- lattice-based cryptography,\\
- random numbers, or\\
- the design and attack of crypto protocols (like SSL).

@@

This document is delivered together with the open-source programs of the
CrypTool project\index{CrypTool}.

@@

Chapters \ref{Chapter_ModernCryptography} (``Modern Cryptography''),
\ref{Chapter_EllipticCurves} (``\nameref{Chapter_EllipticCurves}''),
\ref{Chapter_BitCiphers} (``Bitblock and Bitstream Ciphers''), and % (``\nameref{Chapter_BitCiphers}'')
\ref{Chapter_HomomorphicCiphers} (``\nameref{Chapter_HomomorphicCiphers}'')
require a deeper knowledge in mathematics, while the other chapters should
be understandable with a school leaving certificate.

@@

\noindent \textbf{The programs CrypTool 1\index{CrypTool}\index{CT1},
  CrypTool 2\index{CrypTool 2}\index{CT2} and JCrypTool\index{JCrypTool}\index{JCT}}

CrypTool 1 (CT1) is an educational program with a comprehensive
online help enabling you to use and analyse cryptographic procedures within a
unified graphical user interface. The online help in CrypTool 1 contains both
instructions how to use the program and explanations about the methods itself
(both not as detailled and in another structure than the CT book).

CrypTool 1\index{CrypTool 1} and the successor versions
CrypTool 2 (CT2)\index{CrypTool 2} and JCrypTool (JCT)\index{JCrypTool}
are used world-wide for training in companies and teaching at schools and
universities.
\par \vskip + 15pt

\noindent \textbf{CrypTool-Online\index{CrypTool-Online}\index{CTO}}

Also part of the CT project is the website CrypTool-Online (CTO)
(\url{http://www.cryptool-online.org}), where you can check and apply
cryptographic methods within a browser or on a smartphone. The scope
of CTO is far below from the standalone programs CT1, CT2 and JCT.
\par \vskip + 15pt

@@

MTC3 is the abbreviation for MysteryTwister C3, an international cryptography
contest (\url{http://www.mysterytwisterc3.org}), which is also based on the
CT project.
Here you can find cryptographic riddles in four categories, a high-score list
and a moderated forum. In the meantime more than 4000 active users participate,
and more than 150 challenges are offered (currently 109 of them are solved by
at least one user).xxxxxxxxxxxxxx
\par \vskip + 25pt  % more than +15 to start Sag at the next page

@@

SageMath is a comprehensive open-source CAS package which can be used to easily
program the mathematical methods explained in this book. A speciality of
this CAS is, that the scripting language is Python.\\
SageMath becomes more and more the standard open-source CAS system at universities.
\par \vskip + 15pt

@@

Within this initiative, one and two 2 day courses in cryptology are offered
for pupils and teachers in order to show how attractive MINT subjects like
mathematics, computer science and especially cryptology are.
The course agenda is a virtual secret agent story.\\
In the meantime, these courses took place for several years in Germany in
different towns.\\
All course material is freely available on net
(\url{http://www.cryptool.org/schuelerkrypto/}).\\
All software used is free software (using mostly CT1 and CT2).\\
As all course material is currently available only in German -- we'd be happy
if someone could translate the course material and build an according course
in English.
\par \vskip + 15pt

@@

\par \vskip + 45pt
\noindent Bernhard Esslinger
\par \vskip + 10pt
\noindent Frankfurt/Siegen (Germany), April 2016

>>>>

% $Id$
% ............................................................................
%      P R E F A C E  and  I N T R O D U C T I O N  (Playing together of book and programs)
%
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

\clearpage\phantomsection
\addcontentsline{toc}{chapter}{Preface to the 12th Edition of the CrypTool Book}
\chapter*{Preface to the 12th Edition of the CrypTool Book}

@@

This book was delivered since the year 2000 -- together with the CrypTool~1
(CT1)\index{CT1} package in version 1.2.01. 
Since then it has been expanded and revised in almost every new version of CT1 and CT2.

@@

The later editorial work in LaTeX added footnotes and cross linkages between
different sections, harmonized the index entries, and made some corrections.

%This edition of the CT book again updated some topics:
Compared to edition 11, this edition completely updated the TeX sources
of the document (e.g. one single bibtex file for all chapters and both languages),
and of course, the content of the book was amended, corrected, and
updated with many topics, for instance:
\vspace{-5pt}
\begin{itemize}
%  \item the definitions of the strength of security functions
%        (chapter \ref{cm_Section_Security_Definitions}),
  \item the largest prime numbers (chap. \ref{search_for_very_big_primes}),
%        new factorization records (chap. \ref{RSA-768}),
%  \item progress in cryptanalysis of AES 
%        (chap. \ref{NeueAES-Analyse}) and
%  \item progress in cryptanalysis of hash algorithms 
%        (chap. \ref{collision-attacks-against-sha-1}) and
%  \item progress in ideas for new crypto methods (RSA successor) 
%        (chap. \ref{xxxxxxxxxcm_Brute-force-versus-Symmetr})\index{xxxxxxxxxx} and
  \item the list of movies or novels, in which cryptography or number theory 
        played a major role (see appendix \ref{s:appendix-movies}),
%       and where primes are used as hangers  
%        (see curiouses in \ref{HT-Quaint-curious-Primes-usage}),
  \item the overviews of all functions in
        \hyperlink{appendix-template-overview-CT2}{CrypTool~2 (CT2)},
        \hyperlink{appendix-function-overview-JCT}{JCrypTool (JCT)}, and
        \hyperlink{appendix-function-overview-CTO}{CrypTool-Online (CTO)}
        (see appendix),
  \item further SageMath scripts for cryptography, and the appendix
        \ref{s:appendix-using-sage} about using the computer algebra system
        SageMath,
        % Nguyen and Massierer (see also chap.~\ref{ec:Sage_Massierer}).
  \item the section about the Goldbach conjecture
        (see \ref{L-GoldbachConjecture}) and about twin primes
        (see \ref{L-TwinCousinPrimes}),
  \item the section about shared primes in RSA modules used in reality
        (see \ref{nt_Shared-Primes}),
%  \item the section about RSA fixed points
%        (see \ref{l:NumberTheory_Sage_Number-of-RSA-FixedPoints}), and
%  \item homomorphic encryption
%        (see chapter \ref{Chapter_HomomorphicCiphers}).
  \item the ``\nameref{Chapter_BitCiphers}'' is completely new
        (see chapter \ref{Chapter_BitCiphers}),
  \item the ``\nameref{Chapter_Dlog-FactoringDead}'' is completely new too
        (see chapter \ref{Chapter_Dlog-FactoringDead}). It's a phantastic
	in-depth summary about the limits of the according current
	cryptanalytical methods.
\end{itemize}

@@

At this point I'd like to thank explicitly the following people who
particularly contributed to the CrypTool project\index{CrypTool}. They
applied their very special talents and showed really great engagement:
\vspace{-7pt}
%\begin{itemize}
\begin{list}{\textbullet}{\addtolength{\itemsep}{-0.5\baselineskip}}
   \item Mr.\ Henrik Koy
   \item Mr.\ J\"org-Cornelius Schneider
   \item Mr.\ Florian Marchal
   \item Dr.\ Peer Wichmann
   \item Mr.\ Dominik Schadow
   \item Staff of 
         Prof.\ Johannes Buchmann,
         Prof.\ Claudia Eckert,
         Prof.\ Alexander May,
         Prof.~Torben Weis, and especially
         Prof.~Arno Wacker.
\end{list}

@@

Thanks also to the readers who sent us feedback. And especial thanks for the
free proof reading of this edition done by Helmut Witten and Prof.\ Ralph-Hardo Schulz.
% Especially I would like to acknowledge the English language proof-reading
% of this book version done by Richard Christensen and Lowell Montgomery. Tom.

@@

\par \vskip + 35pt
\noindent Bernhard Esslinger
\par \vskip + 10pt
\noindent Heilbronn/Siegen (Germany), August 2016

\par \vskip + 90pt
\noindent PS:\\
We'd be glad if new authors would show up to improve existing chapters or to add further chapters,
e.g. about
\begin{compactitem}
   \item the Riemann Zeta function,
   \item hash functions and password attacks,
   \item lattice-based cryptography,
   \item random numbers, or
   \item the design and attack of crypto protocols (like SSL).
\end{compactitem}

\par \vskip + 15pt
\noindent PPS:\\%xxxxxxxxxx Todo
Todos to be dealt with to make edition 12 of this CTB a release (till then we still call it a draft):%xxxxxxxxxxxxx
\begin{compactitem}
   \item Update all information about SageMath (chap.~\ref{ec:Sage_Massierer}
      and appendix) and run the code against the newest SageMath (version 7.x) -- 
      both command line and SageMathCloud notebook.
   \item Update the function lists of the four CT versions (in the appendix).
\end{compactitem}

@@

This document is delivered together with the open-source programs of the
CrypTool project\index{CrypTool}. You can also download it directly from the
website of the CT portal: \url{https://www.cryptool.org/en/ctp-documentation}.

@@

Chapters \ref{Chapter_ModernCryptography} (``Modern Cryptography''),
\ref{Chapter_EllipticCurves} (``\nameref{Chapter_EllipticCurves}''),
\ref{Chapter_BitCiphers} (``Bitblock and Bitstream Ciphers''), %% (``\nameref{Chapter_BitCiphers}'')
\ref{Chapter_HomomorphicCiphers} (``\nameref{Chapter_HomomorphicCiphers}''), and
\ref{Chapter_Dlog-FactoringDead} (``Results for Solving Discrete
Logarithms and for Factoring'') % Shortened as original title is too long.
require a deeper knowledge in mathematics, while the other chapters should
be understandable with a school leaving certificate.

@@

\noindent \textbf{The programs CrypTool~1, CrypTool~2, and JCrypTool}
\index{CT1} \index{CT2} \index{JCT}
% \noindent \textbf{The programs CrypTool~1\index{CrypTool~1}\index{CT1},
%   CrypTool~2\index{CrypTool~2}\index{CT2}, and JCrypTool\index{JCrypTool}\index{JCT}}
%%% Solved-Problem__Only in English__Separated index at p144 (4.11.2) and p217 (5.3) and
%%%     at page xviii, but \index{CrypTool~1} was called like everywhere else!
%%%     In *.idx stand statt \indexentry{CrypTool~1|hyperpage}{iii}
%%%     ein                  \indexentry{CrypTool\nobreakspace  {}1|hyperpage}{xviii}
%%%  --> Fixed here by moving the \index out of \textbf.
%%%  --> Fixed in chap 4 (below \label{chptSecurityParam}) by moving the \index out of
%%%      \paragraph (done via moving footnote out of paragraph using footnotemark).
%%%  --> Fixed in 5.3 by moving the \index out of \subsection-Klammern using footnotemark).
%%%      BEFORE_MOVING_OUT_INDEX: \subsection[The RSA procedure]{The RSA procedure\footnote{%
%%%  See: http://wwws.htwk-leipzig.de/~myagovki/latex/fussnoten_und_randnotizen/
%%%       https://en.wikibooks.org/wiki/LaTeX/Counters

CrypTool~1 (CT1) is an educational program enabling you to use and analyze
cryptographic procedures within a unified graphical user interface. The
comprehensive online help in CrypTool~1 contains both
instructions how to use the program and explanations about the methods itself
(both not as detailed and in another structure as in the CT book).

CrypTool~1 and the successor versions CrypTool~2 (CT2) and JCrypTool (JCT)
are used world-wide for training in companies and teaching at schools and
universities.
\par \vskip + 15pt

\noindent \textbf{CrypTool-Online\index{CTO}}

Another part of the CT project is the website CrypTool-Online (CTO)
(\url{http://www.cryptool-online.org}), where you can apply
crypto methods within a browser or on a smartphone. The scope
of CTO is far below from the standalone programs CT1, CT2 and JCT.
However, this is what people more and more use as the first contact,
so we currently redesign the whole backbone system with most modern
web technology to have a fast and consistent look\&feel.
\par \vskip + 15pt

@@

MTC3 is the abbreviation for MysteryTwister C3, an international cryptography
contest (\url{http://www.mysterytwisterc3.org}), which is also based on the
CT project.
Here you can find cryptographic riddles in four categories, a high-score list
and a moderated forum. As of 2016-06-16 more than 7000 users participate,
and more than 200 challenges are offered (162 of them are solved by
at least one participant).
\par \vskip + 25pt  % vskip: more than +15 to start SageMat the next page

@@

SageMath is a comprehensive open-source CAS package which can be used to easily
program the mathematical methods explained in this book. A speciality of
this CAS is, that the scripting language is Python (currently version 2.x).
So in a Sage script, you have after an import statement all functions from the
Python language at your disposal.\\
SageMath becomes more and more the standard CAS system at universities.
\par \vskip + 15pt

@@

Within this initiative, one and two 2 day courses in cryptology are offered
for pupils and teachers in order to show how attractive MINT subjects like
mathematics, computer science and especially cryptology are.
The course agenda is a virtual secret agent story.\\
In the meantime, these courses took place for several years in Germany in
different towns.\\
All course material is freely available at
\url{http://www.cryptool.org/schuelerkrypto/}.\\
All software used is free software (using mostly CT1 and CT2).\\
As all course material is currently available only in German -- we'd be happy
if someone could translate the course material and build an according course
in English.
\par \vskip + 15pt

@@

\par \vskip + 45pt
\noindent Bernhard Esslinger
\par \vskip + 10pt
\noindent Heilbronn/Siegen (Germany), August 2016

### 2016-08-23 script/en/chapters/menus.tex
CT Book ed 12: Added 3 graphics in primes.tex -- at the beginning and into section sieve of Eratosthenes. Started to search for a old /index problem [so later when solved the error detecting patterns ("\index{ZZZ") have to be taken off again.]
CT Book ed 12: Major update -- many changes, corrections, and additions during last 4 months

% $Id%

% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%\pagebreak
\newpage
%\enlargethispage{1cm}
\hypertarget{appendix-menu-overview-CT1}{}
\section{CrypTool 1 Menus}
\label{s:appendix-menu-overview-CT1}

This appendix contains at the following page the complete menu tree of
CrypTool\index{CrypTool} version 1.4.31\footnote{%
  Since 2010, changes for the stable CrypTool 1 ({\bf CT1})\index{CrypTool 1}
  \index{CT1} focus mainly on bugfixes. Many new developments go into the two
  successor versions CrypTool 2 ({\bf CT2})\index{CrypTool 2}\index{CT2} and
  JCrypTool ({\bf JCT})\index{JCrypTool}\index{JCT}:\\
  - Web site CT2: \url{http://www.cryptool.org/en/ct2-documentation-en} \\
  - Web site JCT: \url{http://www.cryptool.org/en/jct-volunteer-en} \\
  These successor versions are currently (Nov. 2012) still beta versions;
  but they are close to their first release and they are stable enough
  since quite a while to be used by end-users.\\
}.

@@

Which of the menu items in CrypTool 1 are active (that means not greyed),
depends on the type of the currently active document window:
The brute-force analysis\index{attack!brute-force} for DES e.~g. is only
available, if the active window is opened in the hexadecimal view. 
On the other hand the menu item ``Generate Random Numbers\dots''
is always available (even if no document is opened).

@@

%--------------------------------------------------------------------
\clearpage
\begin{figure}[hb]
\begin{center}
\vspace{-30pt}

\rotatebox{90}{%
\begin{minipage}{1.57\textwidth}% So sah es für mich mittig aus.
\includegraphics[scale=0.35]{figures/CT1-menutree-en}
\hypertarget{appendix-figure-menu-overview-CT1}{}
\caption{Complete overview of the menu tree of CT1 (CrypTool 1.4.31xxxxxxxxxxxxxxxxxxx)} 
\label{appendix-figure-menu-overview-CT1}
\end{minipage}
}

\end{center}
\end{figure}
\clearpage

%--------------------------------------------------------------------
\newpage
%\enlargethispage{1cm}
\hypertarget{appendix-template-overview-CT2}{}
\section{CrypTool 2 Templates}
\label{s:appendix-template-overview-CT2}

\noindent This appendix contains on the following pages the complete tree with
all templates in CT2\index{CrypTool 2}.\footnote{%
  You can find further information about CT2 at:
  \url{http://www.cryptool.org/en/ct2-documentation-en}
}

@@

\noindent On the following pages this appendix contains a list of all
functions in JCrypTool\index{JCrypTool}.\footnote{%
  You can find further information about JCT at:
  \url{http://www.cryptool.org/en/jct-volunteer-en} \\
  This list was generated using the CT Portal website.}

@@

All functions of the {\bf Default Perspective} can be found both in the menus and in the navigation bar called ``Crypto Explorer'' (at the right side).
The Default Perspective contains all important methods like classic transposition or modern AES, and many visualizations (e.g. Diffie-Hellman key exchange or caculations on elliptic curves).

All functions of the {\bf Algorithm Perspective} can be found in the navigation bar called ``Algorithms'' (in this perspective also at the right side).
The Algorithm Perspective contains all detail settings of the various algorithms, it especially offers post-quantum computing algorithms\index{post-quantum computing}\index{PQC}.

@@

\noindent This appendix contains a list of all
functions in CrypTool-Online (CTO)\index{CrypTool-Online}.\footnote{%
  You can find further information about CTO at:
  \url{www.cryptool-online.org} \\
  This list was generated using the functions list at the CT Portal website:\\
  \url{http://www.cryptool.org/ctp-documentation-en/ctp-functions-en}}

>>>>

% $Id: 
% ..............................................................................
%             M E N U E S  i n  V E R S C H I E D E N E N  C T - V A R I A N T E N
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

\newpage
%\enlargethispage{1cm}
\hypertarget{appendix-menu-overview-CT1}{}
\section{CrypTool~1 Menus}
\label{s:appendix-menu-overview-CT1}

This appendix contains at the following page the complete menu tree of
CrypTool\index{CrypTool} version 1.4.31\footnote{%
  Since 2010, changes for the stable CrypTool~1 ({\bf CT1})\index{CT1}
  focus mainly on bugfixes. Many new developments go into the two
  successor versions CrypTool~2 ({\bf CT2})\index{CT2} and
  JCrypTool ({\bf JCT})\index{JCT}:\\
  - Web site CT2: \url{http://www.cryptool.org/en/ct2-documentation-en} \\
  - Web site JCT: \url{http://www.cryptool.org/en/jct-volunteer-en} \\
  These successor versions are currently (Nov. 2012) still beta versions;
  but they are close to their first release and they are stable enough
  since quite a while to be used by end-users.\\
}.

@@

Which of the menu items in CrypTool~1 are active (that means not greyed),
depends on the type of the currently active document window:
The brute-force analysis\index{attack!brute-force} for DES e.~g. is only
available, if the active window is opened in the hexadecimal view. 
On the other hand the menu item ``Generate Random Numbers\dots''
is always available (even if no document is opened).

@@

%--------------------------------------------------------------------
\clearpage
\begin{figure}[hb]
\begin{center}
\vspace{-30pt}
\rotatebox{90}{%
\begin{minipage}{1.57\textwidth}
\includegraphics[scale=0.35]{figures/CT1-menutree-en}
\hypertarget{appendix-figure-menu-overview-CT1}{}
\caption{Complete overview of the menu tree of CT1 (CrypTool~1.4.31)}%TODO-xxxxxxxxxxxxxxxxx
\label{appendix-figure-menu-overview-CT1}
\end{minipage}
}
\end{center}
\end{figure}
\clearpage

%--------------------------------------------------------------------
\newpage
%\enlargethispage{1cm}
\hypertarget{appendix-template-overview-CT2}{}
\section{CrypTool~2 Templates}
\label{s:appendix-template-overview-CT2}

\noindent This appendix contains on the following pages the complete tree with
all templates in CT2\index{CT2}.\footnote{%
  You can find further information about CT2 at:
  \url{http://www.cryptool.org/en/ct2-documentation-en}
}

@@

\noindent On the following pages this appendix contains a list of all
functions in JCrypTool\index{JCT}.\footnote{%
  You can find further information about JCT at:
  \url{http://www.cryptool.org/en/jct-volunteer-en} \\
  This list was generated using the CT Portal website.}

@@

All functions of the {\bf Default Perspective} can be found both in the menus and
in the navigation bar called ``Crypto Explorer'' (at the right side).
The Default Perspective contains all important methods like classic transposition
or modern AES, and many visualizations (e.g. Diffie-Hellman key exchange or
calculations on elliptic curves).

All functions of the {\bf Algorithm Perspective} can be found in the navigation bar
called ``Algorithms'' (in this perspective also at the right side).
The Algorithm Perspective contains all detail settings of the various algorithms, it
especially offers post-quantum computing algorithms\index{post-quantum computing} (PQC).

@@

\noindent This appendix contains a list of all
functions in CrypTool-Online (CTO)\index{CTO}.\footnote{%
  You can find further information about CTO at:
  \url{www.cryptool-online.org} \\
  This list was generated using the functions list at the CT Portal website:\\
  \url{http://www.cryptool.org/ctp-documentation-en/ctp-functions-en}}
